---
title: "Complex budgets"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Complex budgets}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(om)
```
In the introductory vignette, we saw how Omland could optimise the national budget
given the following modelled cost and impact estimates

```{r}
omland_output <- data.frame(
  unit = rep(1:10, each = 5),
  option = rep(1:5, 10),
  impact = rep(0:4, 10),
  cost = rep(c(0, 5, 10, 15, 20), 10)  * rep(c(rep(1, 5), rep(5, 5)), each = 5)
)
impact_matrix <- matrix(omland_output$impact, nrow = 10, ncol =  5, byrow = TRUE)
cost_matrix <- matrix(omland_output$cost, nrow = 10, ncol =  5, byrow = TRUE)

```

Breaking news!!! A regional initiative has been set up to increase funding in 
two of Omland's sub-national units, units number 9 and 10. An additional $200
has been made available to Omland, but spending must be restricted units 9 and 10.

Fortunately for Omland, om can deal with multiple budget levels. Let's look at how
we can set this up. We now specify two budgets, the national and regional

```{r}
budget <- c(100, 200)
```

We also need to specify which sub-national units have access to which parts of the
budget. We can do this with another matrix. In the `recipients` matrix, each row
is a sub-national unit, and each column corresponds to a level in our budget vector.
A `1` indicates that a sub-national unit may receive a share of that budget, a `0`
indicates it can not receive any of it.

```{r}
recipients <- cbind(
  rep(1, 10),
  c(rep(0, 8), rep(1, 2))
)
recipients
```

Here, we see that only rows 9 and 10 have access to both budget levels. Now we can
optimise again

```{r}
optimised <- om(z = impact_matrix, cost = cost_matrix, budget = budget,
                recipients = recipients)
optimised
```

Now we see that the first budget level, available to all of Omland, is shared
across the first 5 sub-national units, just like in the introductory example. 
However, thanks to the regional initiative, Omland can now spend an additional $200
in sub-national units 9 and 10, improving the overall impact. The columns 
`budget_level_1` and `budget_level_2` show us how each section of Omland's 
budget is distributed. 

You can add more budget levels by modifying the `budget` and `recipients` inputs.

