---
title: "Cost-effectiveness frontier"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Cost-effectiveness frontier}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo=FALSE}
library(om)
```

This vignette demonstrates how to find the cost-effectiveness frontier using
`frontier()`. We'll create a small simulated data set with cost and impact
values and then highlight the solutions that lie on the frontier. We also apply
a *threshold* that limits the maximum cost per unit of impact gained. The World
Health Organization suggests that interventions costing up to 1–3 times a
country's GDP per capita for each disability-adjusted life year (DALY) gained
are often considered cost-effective. In our generic example this simply means
setting an upper limit on the cost divided by impact.

```{r}
set.seed(1)
example <- data.frame(
  cost = runif(50, 1, 10)
)
example$impact <- 15 + example$cost + rnorm(50)

```

We can visualise the frontier relative to all generated points. Points that lie
on the frontier are shown in **pink**. Fully dominated solutions are in **grey**.

```{r, fig.width = 6, fig.height = 5}
frontier_solution <- frontier(example)
plot(
  example$impact ~ example$cost,
  pch = 19,
  xlab = "Cost",
  ylab = "Impact",
  col = "grey50"
  )
points(frontier_solution$impact ~ frontier_solution$cost, col = "deeppink", pch = 19)
legend(
  "bottomright",
  legend = c("Dominated solutions", "Efficient frontier"),
  col = c("grey50", "deeppink"),
  pch = 19,
  bty = "n"
)
```

We can also drop dominant solutions that fall outside of a user-specified
ICER threshold, those filtered out by the threshold are coloured **blue**

```{r, fig.width = 6, fig.height = 5}
frontier_threshold_solution <- frontier(example, threshold = 1)
plot(
  example$impact ~ example$cost,
  pch = 19,
  xlab = "Cost",
  ylab = "Impact",
  col = "grey50"
  )
points(frontier_solution$impact ~ frontier_solution$cost, col = "dodgerblue", pch = 19)
points(frontier_threshold_solution$impact ~ frontier_threshold_solution$cost, col = "deeppink", pch = 19)
legend(
  "bottomright",
  legend = c("Dominated solutions", "Efficient frontier", "Above ICER threshold"),
  col = c("grey50", "deeppink", "dodgerblue"),
  pch = 19,
  bty = "n"
)
```

We can also drop dominant solutions that are not on the convex hull of the frontier.
That leaves solutions shown in **Green**

```{r, fig.width = 6, fig.height = 5}
frontier_threshold_solution_hull <- frontier(example, convex_hull = TRUE, threshold = 1)
plot(
  example$impact ~ example$cost,
  pch = 19,
  xlab = "Cost",
  ylab = "Impact",
  col = "grey50"
  )
points(frontier_solution$impact ~ frontier_solution$cost, col = "dodgerblue", pch = 19)
points(frontier_threshold_solution$impact ~ frontier_threshold_solution$cost, col = "deeppink", pch = 19)
points(frontier_threshold_solution_hull$impact ~ frontier_threshold_solution_hull$cost, col = "green", pch = 19)
legend(
  "bottomright",
  legend = c("Dominated solutions", "Extendedly dominated", "Above ICER threshold", "Efficient frontier (convex hull)"),
  col = c("grey50", "deeppink", "dodgerblue", "green"),
  pch = 19,
  bty = "n"
)
```

The convex frontier is the standard approach in cost-effectiveness analysis: it removes extendedly dominated options so that incremental cost-effectiveness ratios (ICERs) increase monotonically. This identifies the minimal set of strategies that could be optimal when interventions are compared sequentially at a given willingness-to-pay threshold.

The full non-dominated set includes all strategies that are not strictly worse than any other. Although some may be extendedly dominated, this view can be more informative when exploring how the set of cost-effective options might change as budgets increase or decrease.
